---
title: "Shale content methods"

description: |
  Exercise 5.1 Sensitivity on Vsh Calculation 
  Ref. Sanni, M. (2019) Petroleum Engineering. Principles, Calculatios, and Workflows
author:
  - name: Rigoberto Chandomí Vázquez
date: 06-05-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

Gamma Ray logs can be used for determining the shale content in a formation using differents relationship between Vsh and GR. Data.

Vsh Calculation Methods (Sanni, 2019).

|Method| Formula| Formulation and comments|
|:----|:----|:----|
|Linear (Gamma Ray Index)|$$V_{sh} = I_{GR}$$|Based in linear relationship between shale volume and GR response|
|Larionov [1969] for tertiary (unconsolidated) rock|$$V_{sh} = 0.083(2^{3.7I{GR}}-1)$$|Based on empirical correlation. Linear relationship overestimates $V_{sh}$ for tertiary (unconsolidated) rocks.|
|Larionov (1969) for pre-tertiary (older and consolidated) rock|$$V_{sh} = 0.33(2^{2I{GR}}-1)$$|Based on empirical correlation. Linear relationship overestimates $V_{sh}$ pre-tertiary (consolidate) rocks|
|Stieber [1970]|$$V_{sh}=\frac{0.5I_{GR}}{1.5-I_{GR}}$$|Calibration to Gulf coast log|
|Clavier et al. [1971]|$$V_{sh}=1.7-(3.38-(I_{GR}+0.7)^2)^{1/2}$$|Compromise between Larionov Tertiary and old rock model|

$I_{GR}$ is called the gamma ray index and it is defined as:
$$I_{GR} = \frac{GR_{Zone}-GR_{Clean}}{GR_{Shale}-GR_{Clean}}$$

First, read the log data and plot the GR log

```{r, echo=TRUE}
library(ggplot2)
library(dplyr)
library(DT)
library(plotly)

log_data <- read.csv("Ex_5.1_log.csv")

datatable(log_data)

fig <- ggplot(log_data) + geom_line(aes(MD,GR), color = "green", size = 1.25) +
                   coord_flip() + ylab("GR (GAPI)") + xlab("MD (ft)") +
                   scale_x_continuous(trans = "reverse") +
                   scale_y_continuous(position = "right") + ylim(0, 150)

fig

```

According GR plot we can define as $GR_{Shale} = 111.86$ and $GR_{Clean} = 58.12$

```{r, echo=TRUE}

fig <- fig + geom_hline(yintercept = 58.12, linetype="dotted", 
                color = "blue", size=1.5) + 
            geom_hline(yintercept = 111.86, linetype="dotted", 
                color = "blue", size=1.5)

fig

```

Now, we can calculate Gamma Ray Index and Shale content ($V_{sh}$) using the equations above.



```{r, echo=TRUE, fig.height=11}

Vsh <- log_data %>%
        mutate(IGR = (GR-58.12)/(111.86-58.12),
               IGR = ifelse(IGR>1, 1, IGR),
               IGR = ifelse(IGR<0, 0, IGR),
               Vsh_Linear = IGR,
               Vsh_LarionovUn = 0.083*(2^(3.7*IGR)-1),
               Vsh_LarionovC = 0.33*(2^(2*IGR)-1),
               Vsh_Stieber = (0.5*IGR)/(1.5-IGR),
               Vsh_Clavier = 1.7-(3.38-(IGR+0.7)^2)^0.5)  %>%
        select(MD, IGR, Vsh_Linear, Vsh_LarionovUn, Vsh_LarionovC, Vsh_Stieber, Vsh_Clavier)

datatable(Vsh)


plot_Vsh <- ggplot(Vsh) + geom_line(aes(MD,Vsh_Linear, colour = "Linear"), size = 1.25) +
            geom_line(aes(MD,Vsh_LarionovUn, colour = "Larinov (Tertiary)"), linetype = "dashed", size = 1.25) +
            geom_line(aes(MD,Vsh_LarionovC, colour="Larinov (Pre-Tertiary)"), linetype = "dotted", size = 1.25) +
            geom_line(aes(MD,Vsh_Stieber, colour = "Steiber"), linetype = "longdash", size = 1.25) +
            geom_line(aes(MD,Vsh_Clavier, colour = "Clavier"), linetype = "dotdash", size = 1.25) +
            coord_flip() + ylab("Vsh") +
            scale_x_continuous(trans = "reverse") +
            theme(text = element_text(size=14), legend.position="right") +
            scale_color_manual(name = "Vsh methods", breaks = c("Linear", "Larinov (Tertiary)",
                                                                  "Larinov (Pre-Tertiary)",
                                                                  "Steiber","Clavier"),
                                        values = c("black", "red" , "green4", "blue","brown"))

plot_Vsh
#ggplotly(plot_Vsh)

```

References 
Sanni, M. (2019) Petroleum Engineering. Principles, Calculatios, and Workflows

